Off the beaten path: \\ Gaia reveals GD-1 stars perturbed off of the main stream track

### 2018-04-29 paper/gd1.tex
MAD
abstract

\begin{abstract}\noindent % trust me
Tidally-disrupted globular clusters are transformed into thin, dynamically-cold
streams of stars that are extremely valuable tracers of the large- and
small-scale distribution of mass in the Galaxy.
Most stellar streams discovered around the Milky Way reside in the Galactic halo
and are therefore primarily sensitive to the dark matter distribution.
We present a sample of highly-probable members of the longest cold stream, GD-1,
selected to have small parallaxes and retrograde proper motions using data from
the \gaia\ mission, and \pans\ photometry consistent with an old and metal-poor
population at a distance of $\sim10~\kpc$.
This selection extended the stream by $20^\circ$, revealed a possible location of the progenitor and mapped significant density variations along the stream in unprecedented detail.
In addition to several prominent gaps, for the first time we also find evidence of stream members off the main track.
No(?) secular formation mechanisms predict the existence of such spurs, but they are a natural result of a stream interaction with a massive perturber.
Using the impulse approximation arguments, we estimate that the detected underdensity and the associated spur could have been induced by a perturbation of a $\sim10^7\,\rm M_\odot$(?) object $\sim 0.5\,$Gyr ago.
Our model of GD-1 indicates that the stream crosses the Galactic plane at a large radius, which puts the likely origin of the perturber beyond the stellar disk.
\end{abstract}

@@

We extract properties of the stream from the stream region defined above as a
function of $\phi_1$ by computing properties in overlapping windows:
we set the window size to $XX^\circ$ in $\phi_1$, and shift the window center by
$1^\circ$ along the stream region.
\figurename~\ref{fig:track-and-model} shows the median latitude, $\phi_2$,
stream surface density, $XX$, stream width, $XX$, and median proper motions,
$\mu_{\phi_1}$ and $\mu_{\phi_2}$, extracted in this way.
For the surface density, we subtract an estimate of the background surface
density by \todo{AB}.
To estimate the stream width, $w$, we compute the median absolute deviation,
$\textrm{MAD}$, and use this to estimate a robust standard deviation, $w = 1.5
\times MAD$.

>>>>

\begin{abstract}\noindent % trust me
Tidally-disrupted globular clusters are transformed into thin, dynamically-cold
streams of stars that are extremely valuable tracers of the large- and
small-scale distribution of mass in the Galaxy.
Most stellar streams discovered around the Milky Way reside in the Galactic halo
and are therefore primarily sensitive to the dark matter distribution.
The longest and most prominent thin stream --- the GD-1 stream --- has been used
to constrain the shape of the inner halo, and is a promising place to search for
signatures of dark matter subhalo interactions.
% Aims
Here we present a sample of highly-probable members of GD-1, selected to have
small parallaxes and retrograde proper motions using data from the \gaia\
mission, and \pans\ photometry consistent with an old and metal-poor population
at a distance of $\sim10~\kpc$.
This selection extends the stream by $20^\circ$, reveals a plausible location of
the progenitor, and shows significant density variations along the stream in
unprecedented detail.
In addition to several prominent gaps, for the first time we also find evidence
of stream members off the main track.
Such features are a natural result of a stream interaction with a massive
perturber, or an indication of complex internal kinematics within the
stream progenitor.
If caused by a subhalo interaction, we estimate that the detected underdensity
and the associated spur could have been induced by a perturbation of a
$\sim10^7\,\rm M_\odot$(?) object $\sim 0.5\,$Gyr ago.
\end{abstract}

@@

We extract properties of the stream from the stream region defined above as a
function of $\phi_1$ by computing properties in overlapping windows:
we set the window size to $XX^\circ$ in $\phi_1$, and shift the window center by
$1^\circ$ along the stream region.
\figurename~\ref{fig:track-and-model} shows the median latitude, $\phi_2$,
stream surface density, $XX$, stream width, $XX$, and median proper motions,
$\mu_{\phi_1}$ and $\mu_{\phi_2}$, extracted in this way.
For the surface density, we subtract an estimate of the background surface
density by \todo{AB}.
To estimate the stream width, $w$, we compute the median absolute deviation,
$\textrm{MAD}$, and use this to estimate a robust standard deviation, $w = 1.5
\times \textrm{MAD}$.
